---
title: "flchain data analysis"
author: "David Rossell and Francisco Rubio"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    df_print: paged
---


We load R packages and a file `functions.R` containing auxiliary functions.

```{r,message=FALSE}
library(survival)
library(devtools)
install_github("davidrusi/mombf")
library(mombf)
library(Biobase)
package.version("mombf")
```

We load the `flchain` data set from the R package `survival`. This data set contains a stratified random sample containing half of the subjects from a study of the relationship between serum free light chain (FLC) and mortality. Only complete cases are analyzed.
The response variable is the number of days from enrollment until death. Note that there are 3 subjects whose sample was obtained on their death date (`futime`).
We consider the following 5 covariates: `sex` (0 - Female, 1 - Male), serum free light chain kappa portion (`kappa`), serum free light chain lambda portion (`lambda`), serum creatinine (`creatinine`), diagnosed with monoclonal gammapothy (`mgus`, 0-1).
The variable `age` is left-truncated at age 50. Thus, we do not include this variable in our selection as this dataset only includes persons from an advanced age group.
For further information type `help(flchain)`.
We scale the response variable by 365.25, so that survival times are measured in years, and the continuous covariates to zero mean and unit variance.

```{r}
# flchain data set
data(flchain)
head(flchain)
# Complete cases of the variables of interest
data.c <- flchain[,c(2,4,5,7,8,9,10)]
data.c <- data.c[complete.cases(data.c),]
dim(data.c)
head(data.c)
# Assiging half-day survival to the zero-survivors
data.c$futime <- ifelse(data.c$futime==0,0.5,data.c$futime)
y= Surv(log(data.c$futime/365.25), data.c$death)
data.c$kappa <- scale(data.c$kappa) # scaled kappa
data.c$creatinine <- scale(data.c$creatinine) # scaled creatinine 
data.c$sex <- as.numeric(data.c$sex)-1 # sex 0-1
X= data.c[,-c(6,7)];  # Design matrix
dim(X)
```

The dataset has $n=6,524$ observations, $n_o=1,962$ individuals died during the follow-up period while $n_c=4,562$ where alive at the last contact date. The following plot shows the Kaplan-Meier estimator of the survival function.

```{r}
km <- survfit(Surv(futime/365.25, death) ~ 1, data = data.c)
plot(km, xlab = "Years", ylab = "Survival", main = "Kaplan-Meier estimator",
     cex.axis = 1.5, cex.lab = 1.5)
```

# BAYESIAN MODEL SELECTION

## Linear regression
In this section, we run Bayesian model selection under the block Zellner's priors considering only linear effects (the dispersion is parameterized as tau=$n g_N$, hence the default $g_N=1$ corresponds to tau=$n$) and a Beta-Binomial(1,1) prior on the models. We use option enumerate=TRUE to force full model enumeration (alternatively it uses Gibbs sampling, the obtained results are very similar either way). The task finishes quickly, in less than 1 second. 
```{r}
# Regression formula with linear effects
f1= ~  sex + kappa + lambda + creatinine +  mgus 
f1y= y ~ sex + kappa + lambda + creatinine +  mgus 
X= model.matrix(f1, data=data.c)

# BMS
priorCoef= zellnerprior(tau=nrow(data.c))
priorGroup= zellnerprior(tau=nrow(data.c))
priorDelta= modelbbprior(1,1)
ms.z <- modelSelection(f1y,data=data.c,
                       priorCoef=priorCoef,priorGroup=priorGroup,priorDelta=priorDelta,
                       enumerate=TRUE, method='Laplace')
```

The top model has posterior probability 0.69 and it includes 2 variables (`kappa` and `lambda`), with marginal posterior inclusion probabilities 1 and 0.833 respectively. This is as predicted by theory, when $n$ is large posterior probabilities should concentrate on the optimal model. The model identifier indicates what columns in $X$ are selected.

```{r}
pp.z= postProb(ms.z)
head(pp.z)
```


```{r}
ms.z$margpp
```

## Non-linear regression
In this section, we consider BMS with non-linear effects over all continuous variables and under block Zellner's priors. 
```{r}
priorCoef= zellnerprior(tau=nrow(X))
priorGroup= groupzellnerprior(tau=nrow(X))
priorDelta= modelbbprior(1,1)
system.time(ms.zn <- modelSelection(y~ sex + kappa + lambda + creatinine + mgus,
                                    smooth=~ kappa + lambda + creatinine, data=data.c, 
                                    priorCoef=priorCoef,priorGroup=priorGroup,priorDelta=priorDelta,
                                    enumerate=TRUE, method='Laplace'))
```

The top model contains again kappa and lambda, but also a non-linear effect of kappa (columns labelled kappa.s, for "smooth"), a non-linear effect of creatinine that was missed by the linear model (columns creatinine.s), and the variable gender. The marginal inclusion probabilities are all $>0.95$, except for gender which is close to 0.5.
```{r}
pp.zn= postProb(ms.zn)
head(pp.zn)
```


```{r}
ms.zn$margpp
```


# COMPARISON TO FREQUENTIST AFT MODELS
We fit Accelerated Failure Time (AFT) models with `lognormal` baseline hazard, via maximum likelihood estimation, for the model with all variables (linear effects), as well as the top models obtained from selecting (i) linear effects, (ii) non-linear effects (using the command `bs()` from the R package `splines`). We note that AIC favours the model containing non-linear effects.
```{r}
library(splines)
# AFT model using a lognormal baseline distribution and all variables (linear model)
fit.aft <- survreg(formula = Surv(futime/365.25, death) ~ sex + kappa + lambda + creatinine + mgus, 
                     data = data.c, dist = "lognormal")

summary(fit.aft)
```


```{r}
# AFT model using a lognormal baseline distribution and selected variables (linear model)
fit.aft.l <- survreg(formula = Surv(futime/365.25, death) ~  kappa + lambda, 
                     data = data.c, dist = "lognormal")

summary(fit.aft.l)
```


```{r}
fit.aft.n <- survreg(formula = Surv(futime/365.25, death) ~ lambda + bs(kappa,df=5) + bs(creatinine,df=5), 
                     data = data.c, dist = "lognormal")

summary(fit.aft.n)
```


```{r}
# AIC
AIC(fit.aft)
AIC(fit.aft.l)
AIC(fit.aft.n)
```


## ANALYSIS OF THE ESTIMATED MEDIAN SURVIVAL TIMES 
We now compare the estimated survival under the models selected when considering linear and non-linear effects. 
```{r} 
# Median survival times obtained with the linear model 
predl <- predict(fit.aft.l, type='quantile', p=0.25) 

# Median survival times obtained with the non-linear model 
predn <- predict(fit.aft.n, type='quantile', p=0.25)
```

There are non-negligible differences between these estimates, that is considering non-linear effects had a practical impact.
```{r} 
plot(predl, predn, xlab = "Predicted 0.75 survival quantile in years (linear model)", ylab = "Predicted 0.75 survival quantile in years (non-linear model)", 
cex.axis = 1.5, cex.lab = 1.4, pch='.')
abline(a = 0, b= 1, lwd = 2)
```

As further illustration we plot the average estimated 0.75 survival quantile as a function of kappa under the linear and non-linear models. The non-linear model suggests a stronger effect than the linear model, in the sense that for small values of kappa it predicts a relatively higher survival, whereas for large values of kappe it predicts lower survival.

```{r}
qkappa= quantile(data.c$kappa, probs=c(.1,.9))
kappar= cut(data.c$kappa, breaks= c(-Inf,seq(qkappa[1],qkappa[2],length=10),Inf))
meankappa= tapply(data.c$kappa, INDEX=kappar, FUN=mean)
meanpredn= tapply(predn, INDEX=kappar, FUN=mean)
meanpredl= tapply(predl, INDEX=kappar, FUN=mean)
plot(meankappa, meanpredn, type='l', xlab='kappa', ylab='Predicted survival (0.25 quantile)', cex.axis=1.4, cex.lab=1.4)
lines(meankappa, meanpredl, col='black', lty = 2, lwd=2)
legend('topright', c('Linear','Non-linear'), col=c('black','black'), lwd=c(2,2), lty = c(1,2), cex=1.4)
```

