---
title: "GGM Simulation. Transfer learning"
output: 
  html_document:
    toc: true
    theme: united
---


```{r, warning=FALSE, message=FALSE}
library(modelSelection)
library(mvtnorm)
library(tidyverse)
library(lamW)      # Lambert's W_q function
set.seed(20260710) # for reproducibility
save2pdf <- TRUE
nsims <- 20 # number of simulations
nseq <- c(10, 20, 100) # sample sizes
kseq <- c(20, 50) # problem dimensions
names(kseq) <- paste('k=', kseq, sep='')
```

# Set default prior parameters

We consider a discrete spike-and-slab prior on the precision matrix, with a Beta-Binomial(1,2) prior. That is, the edge inclusion indicators are $Z_{ij} \sim \mbox{Bern}(\eta)$ with $\eta \sim \mbox{Beta}(5,95)$, with prior mean $E(\eta)= 5/100$.
The prior on diagonal elements is $\theta_{ii} \sim \mbox{Exp}(\lambda/2)$, with expectation $\mu= 2/\lambda$, and the prior on off-diagonal elements is $\theta_{ij} \sim N(0, \sigma^2)$. A natural default is $\sigma^2 =1$. If all the columns in the observed dataset are standardized to zero sample mean and unit sample variance, this default is related to Zellner's unit information prior in regression.
Out of the following three settings, we will focus only on Settings 1 and 3.

1. A default $\lambda=1$, and $\sigma^2=1$.

2. A default $\lambda=1$, with $\sigma^2$ specified such that $P(\Theta \succ 0) = 0.95$. This guarantees that the total variation distance between the priors before and after positive-definiteness truncation is small, and hence that the prior is interpretable.

3. Set $\lambda$ such that $P(\Theta_{ii} \succ 0 \mid Z) \geq 1 - \alpha$ for every edge configuration $Z$ and taking $\alpha=0.1$, and $\sigma^2=1$.


```{r}
edge_prob_succ= 5 # prior number of successes 
edge_prob_fail= 95 # prior number of failures
mu = 9.5       # mean of off-diagonal 
lambda= 2/mu
```

We next obtain the value of $\lambda$ for Setting 3.

```{r}
sd.slab <- 1
alpha <- 0.1
x <- - alpha^2 /(32 * exp(2) * kseq^2)
l <- lambertWm1(x) # q = -1 branch of Lambert's W_q function
mu <- (2 * sd.slab * kseq^(3/2) / alpha) * sqrt(-l)
lambda2 <- 2/mu
names(lambda2) <- names(kseq)
lambda2
```

We store the prior parameter values obtained for each setting and $k = \mbox{dim}(\Theta)$.

```{r}
priorpars <- expand.grid(setting=c('Setting 1', 'Setting 3'), k=kseq)
priorpars$edge_prob_succ <- edge_prob_succ
priorpars$edge_prob_fail <- edge_prob_fail
priorpars$lambda <- priorpars$var.slab <- NA
for (i in 1:nrow(priorpars)) {
  if (priorpars$setting[i] == 'Setting 1') {
    priorpars$lambda[i] <- 1
    priorpars$var.slab[i] <- 1
  } else if (priorpars$setting[i] == 'Setting 3') {
    priorpars$lambda[i] <- lambda2[paste('k=', priorpars$k[i], sep='')]
    priorpars$var.slab[i] <- 1
  }
}
priorpars
```


# Prior distribution on model size

We first plot the prior on the model size before truncation.

```{r}
dbetabin <- function(x, n, alpha, beta) {
  ans <- lchoose(n, x) + lgamma(x + alpha) + lgamma(n - x + beta) - lgamma(n + alpha + beta) - lgamma(alpha) - lgamma(beta) + lgamma(alpha + beta)
  exp(ans)
}
```

```{r, out.width="50%"}
for (i in 1:length(kseq)) {
  maxedges <- (kseq[i] * (kseq[i] - 1) / 2)
  plot(0:maxedges, dbetabin(0:maxedges, n=maxedges, edge_prob_succ, edge_prob_fail), type='l', ylab='Prior (before trunc)', xlab="number of edges", main=paste0('k=',kseq[i]))
  abline(v= maxedges * edge_prob_succ / (edge_prob_succ + edge_prob_fail))
}
```

Now, the prior on the model size after the +def truncation.

```{r}
niter <- 50000
priorModel <- modelbbprior(edge_prob_succ, edge_prob_fail)
priorCoef <- normalidprior(1)
priorDiag <- exponentialprior(1)
simprior <- vector("list", length(kseq))
for (i in 1:length(kseq)) {
  y <- matrix(nrow=0, ncol=kseq[i])
  simprior[[i]] <- modelSelectionGGM(y, priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel, niter=niter, verbose=TRUE)
}
```

```{r}
total_edges <- vector("list", length(kseq))
for (i in 1:length(kseq)) {
  diagentry <- simprior[[i]]$indexes[1,] == simprior[[i]]$indexes[2,]
  total_edges[[i]] <- rowSums(as.matrix(simprior[[i]]$postSample[,!diagentry] != 0))
}

nedges <- data.frame(number_edges = unlist(total_edges), k= rep(kseq, sapply(total_edges, length)))
```

```{r, message=FALSE}
ggplot(nedges, aes(x=number_edges)) +
  geom_histogram(aes(y = after_stat(density)), position="identity") +
  facet_wrap(~ k, labeller = labeller(k = function(x) paste0("k = ", x))) +
  labs(x = "Number of edges", y = "Prior probability") +
  theme_bw(base_size = 18) +
  scale_color_grey()
```



# Function to run simulations

We create a function that simulates data from a multivariate Gaussian, fits a Bayesian Gaussian graphical model model under a discrete spike-and-slab prior, and record the mean squared estimation error, the average estimate of each $\theta_{ij}$ entry, and the average posterior probability $P(\theta_{ij} \neq 0 \mid y)$.


```{r}
run_simulations <- function(n, nsims, Theta, Sigma=solve(Theta), priorCoef, priorDiag, priorModel, scale=TRUE) {
  posterior_coef = posterior_nbedges = vector("list", nsims)
  for (i in 1:nsims) {
    set.seed(i)  #for reproducibility

    ansi = ansi_nbedges = vector("list", length(n))
    for (j in 1:length(n)) {
    
      # Simulate data
      samplesize <- n[j]
      y = rmvnorm(samplesize, sigma= Sigma)
      sy = apply(y, 2, 'sd')

      # Fit models   
      fit = modelSelectionGGM(y=y, priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel, center=scale, scale=scale, verbose=FALSE)
      
      # Posterior distribution on the number of edges
      diagentry <- fit$indexes[1,] == fit$indexes[2,]
      ansi_nbedges[[j]] <- data.frame(number_edges = rowSums(as.matrix(fit$postSample[,!diagentry] != 0)), n = samplesize)

      # Extract BMA parameter estimates and posterior edge inclusion probabilities
      b= data.frame(coef(fit)) |>
        select(row, column, estimate, margpp)

      # Convert to original scale
      scaling = 1 / (sy[b$row] * sy[b$column])
      b = transform(b, estimate= estimate * scaling)

      # Add data-generating truth
      if (i==1) {
        truth = double(nrow(b))
        for (l in 1:nrow(b)) truth[l] = Theta[b$row[l], b$column[l]]
      }
      b$truth = truth
  
      # Record MSE, average posterior inclusion probability and average point estimate
      ansi[[j]] = group_by(b, truth) |>
                 summarize(mse = mean((truth - estimate)^2), pp = mean(margpp), mean_estimate = mean(estimate), count = n()) |>
                 cbind(simid = i, n = samplesize)

      cat('.')

    }
    
    posterior_coef[[i]] <- do.call(rbind, ansi)
    posterior_nbedges[[i]] <- do.call(rbind, ansi_nbedges)
  }
  ans = list(posterior_coef=posterior_coef, posterior_nbedges=posterior_nbedges)
  return(ans)
}
```



# Erdos-Renyi graph

## Run simulations


```{r}
true_edge_prob <- edge_prob_succ / (edge_prob_succ + edge_prob_fail)
sims1 <- sims3 <- Theta_truth <- vector("list", length(kseq))
names(sims1) <- names(sims3) <- names(kseq)
diagvals <- vector("list", length(kseq))
for (i in 1:length(kseq)) {
  ksel <- kseq[i]  # number of variables
  cat("Simulations for k =", ksel)
  
  # Data-generating precision matrix
  y <- matrix(nrow=0, ncol=ksel)
  sample <- modelSelectionGGM(y, priorCoef=normalidprior(1), priorModel=modelbinomprior(true_edge_prob), priorDiag=exponentialprior(1), niter=5000, verbose=FALSE)  # simulate from prior
  
  diagentry <- sample$indexes[1,] == sample$indexes[2,]
  sample_edges <- rowSums(as.matrix(sample$postSample[,!diagentry] != 0))
  diff_edges <- abs(sample_edges - true_edge_prob * ksel * (ksel - 1) / 2)
  sel <- rev(which(diff_edges == min(diff_edges)))[1]  # take draw closest to true edge prob
  Theta <- diag(ksel)
  thetavals <- sample$postSample[sel,]
  for (j in 1:ncol(sample$indexes)) {
    Theta[sample$indexes['row',j], sample$indexes['column',j]] <- thetavals[j]
    Theta[sample$indexes['column',j], sample$indexes['row',j]] <- thetavals[j]
  }
  Sigma <- solve(Theta)
  diagvals[[i]] <- unique(diag(Theta))
  Theta_truth[[i]] <- Theta

  # Set prior distribution under each setting
  pars_sel <- filter(priorpars, setting == "Setting 1", k == ksel)
  priorModel <- modelbbprior(pars_sel$edge_prob_succ, pars_sel$edge_prob_fail)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims1[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel) 

  pars_sel <- filter(priorpars, setting == "Setting 3", k == ksel)
  priorModel <- modelbbprior(pars_sel$edge_prob_succ, pars_sel$edge_prob_fail)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims3[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel) 

  cat("\n")
}
diagvals <- unique(unlist(diagvals))
```


We summarize the results across all simulations. For each prior parameter setting and $k= \mbox{dim}(\Theta)$, and each value of the data-generating $\Theta_{ij}$, we report the mean squared estimation error in the Bayesian model averaging estimate $E(\Theta_{ij} \mid y)$, the average marginal posterior edge inclusion probability $P(\Theta_{ij} \neq 0 \mid y)$, and the average BMA estimate.

```{r}
summarize_simulations <- function(kseq, sims1, sims2, sims3) {
  
  sim_summary <- vector("list", length(kseq))
  names(sim_summary) <- names(kseq)

  for (i in 1:length(kseq)) {

    summary1 = do.call(rbind, sims1[[i]]) |>
      data.frame() |>
      group_by(n, truth) |>
      summarize(across(mse:count, mean), .groups="drop") |>
      transform(k = kseq[i], Setting = "Setting 1")

    if (!missing(sims2)) {
      summary2 = do.call(rbind, sims2[[i]]) |>
        data.frame() |>
        group_by(n, truth) |>
        summarize(across(mse:count, mean), .groups="drop") |>
        transform(k = kseq[i], Setting = "Setting 2")
    } else {
      summary2 = NULL
    }
    
    if (!missing(sims3)) {
      summary3 = do.call(rbind, sims3[[i]]) |>
        data.frame() |>
        group_by(n, truth) |>
        summarize(across(mse:count, mean), .groups="drop") |>
        transform(k = kseq[i], Setting = "Setting 3")
    } else {
      summary3 = NULL
    }
  
    sim_summary[[i]] <- rbind(summary1, summary2, summary3)

  }

  sim_summary <- do.call(rbind, sim_summary) |>
    transform(root_mse = round(sqrt(mse),3), pp = round(pp,3), mean_estimate = round(mean_estimate, 4)) |>
    select(-mse) |>
    select(Setting, k, root_mse, everything())

  rownames(sim_summary) <- NULL

  return(sim_summary)
}

```


We summarize the results across all simulations. For each prior parameter setting and $k= \mbox{dim}(\Theta)$, and each value of the data-generating $\Theta_{ij}$, we report the mean squared estimation error in the Bayesian model averaging estimate $E(\Theta_{ij} \mid y)$, the average marginal posterior edge inclusion probability $P(\Theta_{ij} \neq 0 \mid y)$, and the average BMA estimate.

```{r, warning=FALSE}
sims1_coef <- lapply(sims1, function(z) z$posterior_coef)
sims3_coef <- lapply(sims3, function(z) z$posterior_coef)
sim_summary <- summarize_simulations(kseq=kseq, sims1=sims1_coef, sims3=sims3_coef) |>
  filter(!(truth %in% diagvals))

rmse_total <- mutate(sim_summary, weighted_mse = root_mse^2 * count) |>
  group_by(Setting, k, n) |>
  summarize(rmse_total = sqrt(sum(weighted_mse)))

# To facilite plotting, group data-generating truth into bins
# (small bin around 0, to assess post prob to true zeroes)
breaks <- c(min(sim_summary$truth)-0.001,-2,-1,-0.01,0.01,1,2,max(sim_summary$truth)+0.001)
labels <- round((breaks[-1] + breaks[-length(breaks)])/2, 2)
sim_summary <- transform(sim_summary, truth = cut(truth, breaks=breaks, labels=labels)) |>
  group_by(Setting, k, n, truth) |>
  summarize(root_mse = mean(root_mse), pp = mean(pp), mean_estimate=mean(mean_estimate), .groups = "drop") |>
  transform(truth = as.numeric(as.character(truth)))
```


## Plots

Average posterior on the model size.

```{r}
sims1_nbedges <- lapply(sims1, function(z) z$posterior_nbedges)
sims1_nbedges <- lapply(sims1_nbedges, function(z) do.call(rbind, z))
k_column <- rep(names(sims1_nbedges), sapply(sims1_nbedges, nrow))
sims1_nbedges <- do.call(rbind, sims1_nbedges) |>
  mutate(k = k_column)

sims3_nbedges <- lapply(sims3, function(z) z$posterior_nbedges)
sims3_nbedges <- lapply(sims3_nbedges, function(z) do.call(rbind, z))
k_column <- rep(names(sims3_nbedges), sapply(sims3_nbedges, nrow))
sims3_nbedges <- do.call(rbind, sims3_nbedges) |>
  mutate(k = k_column)

setting_column <- rep(c('Setting 1','Setting 3'), c(nrow(sims1_nbedges), nrow(sims3_nbedges)))
sims_nbedges <- rbind(sims1_nbedges, sims3_nbedges) |>
  mutate(Setting = setting_column)
rownames(sims_nbedges) <- NULL
```

```{r, message=FALSE}
vline_data <- data.frame(k=names(kseq), true_nbedges= sapply(Theta_truth, function(z) sum(z[upper.tri(z)] != 0)))

ggplot(sims_nbedges, aes(x=number_edges, col=Setting)) +
  geom_histogram(aes(y = after_stat(density)), fill=NA, position="identity") +
  geom_vline(data = vline_data, aes(xintercept = true_nbedges)) +
  facet_grid(rows= vars(n), cols= vars(k), scale="free_x", labeller=labeller(n = function(x) paste0("n = ", x))) +
  theme_bw(base_size = 18) +
  theme(legend.position = c(0.8, 0.88), legend.title = element_blank()) +
  scale_color_grey() +
  xlab("Number of edges") +
  ylab("Posterior probability")

fname <- "sim_transferlearn_modelsize.pdf"
if (save2pdf) ggsave(fname)
```

We show the average posterior marginal inclusion probability $P(\Theta_{ij} \neq 0 \mid y)$ for prior settings 1 and 3, separately for each $(k,n)$ combination.

```{r, message=FALSE}
filter(sim_summary, !(truth %in% diagvals)) |>
ggplot(aes(x=truth, y=pp, col=Setting, lty=Setting, pch=Setting)) +
  geom_line() +
  geom_point() +
  facet_grid(rows=vars(n), cols=vars(k), 
             labeller = labeller(n = function(x) paste0("n = ", x), k = function(x) paste0("k = ", x))) +
  theme_bw(base_size = 15) +
  theme(legend.position = c(0.9, 0.15), legend.title=element_blank()) +
  scale_color_grey() +
  xlab(expression("Data-generating " * Theta[ij])) +
  ylab("Average posterior inclusion probability")

fname <- "sim_transferlearn_pp.pdf"
if (save2pdf) ggsave(fname)
```



We display the root mean square error, and subsequently the average posterior inclusion probability.

```{r, message=FALSE}
filter(sim_summary, !(truth %in% diagvals)) |>
ggplot(aes(x=truth, y=root_mse, col=Setting, lty=Setting, pch=Setting)) +
  geom_line() +
  geom_point() +
  facet_grid(rows=vars(n), cols=vars(k), 
             labeller = labeller(n = function(x) paste0("n = ", x), k = function(x) paste0("k = ", x))) +
  theme_bw(base_size = 15) +
  theme(legend.position = c(0.1, 0.1), legend.title=element_blank()) +
  scale_color_grey() +
  xlab(expression("Data-generating " * Theta[ij])) +
  ylab("Root mean squared error")

fname <- "sim_transferlearn_rootmse_nodiag.pdf"
if (save2pdf) ggsave(fname)
```



