---
title: "GGM Simulation. Effect of prior slab variance"
output: 
  html_document:
    toc: true
    theme: united
---


```{r, warning=FALSE, message=FALSE}
library(modelSelection)
library(mvtnorm)
library(tidyverse)
library(lamW)      # Lambert's W_q function
set.seed(20260710) # for reproducibility
save2pdf <- TRUE
nsims <- 20 # number of simulations
nseq <- c(100, 500)    # sample size
```

# Set default prior parameters

We consider a discrete spike-and-slab prior on the precision matrix, with prior slab (edge inclusion) probability $\eta=0.1$. 
The prior on diagonal elements is $\theta_{ii} \sim \mbox{Exp}(\lambda/2)$, with expectation $\mu= 2/\lambda$, and the prior on off-diagonal elements is $\theta_{ij} \sim N(0, \sigma^2)$. A natural default is $\sigma^2 =1$. If all the columns in the observed dataset are standardized to zero sample mean and unit sample variance, this default is related to Zellner's unit information prior in regression.
We consider three settings:

1. A default $\lambda=1$, and $\sigma^2=1$.

2. A default $\lambda=1$, with $\sigma^2$ specified such that $P(\Theta \succ 0) = 0.95$. This guarantees that the total variation distance between the priors before and after positive-definiteness truncation is small, and hence that the prior is interpretable.

3. Set $\lambda$ such that $P(\Theta_{ii} \succ 0 \mid Z) \geq 1 - \alpha$ for every edge configuration $Z$ and taking $\alpha=0.1$, and $\sigma^2=1$.


```{r}
edge_prob= 0.1 # probability of non-zero entry
mu = 9.5       # mean of off-diagonal 
lambda= 2/mu
pexp(1, rate=lambda/2, lower.tail=FALSE) #equals 0.9
```

We next obtain the value of $\lambda$ for Setting 3.

```{r}
sd.slab <- 1
alpha <- 0.1
kseq <- c(50, 100)
names(kseq) <- paste('k=', kseq, sep='')
x <- - alpha^2 /(32 * exp(2) * kseq^2)
l <- lambertWm1(x) # q = -1 branch of Lambert's W_q function
mu <- (2 * sd.slab * kseq / alpha) * sqrt(-l)  # liberal upper bound
#mu <- (2 * sd.slab * kseq^(3/2) / alpha) * sqrt(-l) # conservative upper-bound (same results)
lambda2 <- 2/mu
names(lambda2) <- names(kseq)
lambda2
```

Finally, for setting 2, given the prior edge inclusion probability $\eta$ and $\lambda$, function `randompdmatrix_setpars` finds the value of $\sigma^2$ such that $P(\Theta \succ 0)=0.95$. This is done via Monte Carlo, that is sampling from the prior and finding $\sigma$ such that $P(\Theta \succ 0) = 0.95$.

```{r}
var.slab.setting2 <- double(length(kseq))
names(var.slab.setting2) <- paste('k=', kseq, sep='')
for (i in 1:length(kseq)) {
  var.slab.setting2[i] <- modelSelection:::randompdmatrix_setpars(p=kseq[i], edge_prob=edge_prob, diag_distrib='Exp', diag_mean=2, method='MC')$offdiag_variance
}
```

We store the prior parameter values obtained for each setting and $k = \mbox{dim}(\Theta)$.

```{r}
priorpars <- expand.grid(setting=c('Setting 1', 'Setting 2', 'Setting 3'), k=kseq)
priorpars$edge_prob <- edge_prob
priorpars$lambda <- priorpars$var.slab <- NA
for (i in 1:nrow(priorpars)) {
  if (priorpars$setting[i] == 'Setting 1') {
    priorpars$lambda[i] <- 1
    priorpars$var.slab[i] <- 1
  } else if (priorpars$setting[i] == 'Setting 2') {
    priorpars$lambda[i] <- 1
    priorpars$var.slab[i] <- var.slab.setting2[paste('k=', priorpars$k[i], sep='')]
  } else if (priorpars$setting[i] == 'Setting 3') {
    priorpars$lambda[i] <- lambda2[paste('k=', priorpars$k[i], sep='')]
    priorpars$var.slab[i] <- 1
  }
}
priorpars
```



# Function to run simulations

We create a function that simulates data from a multivariate Gaussian, fits a Bayesian Gaussian graphical model model under a discrete spike-and-slab prior, and record the mean squared estimation error, the average estimate of each $\theta_{ij}$ entry, and the average posterior probability $P(\theta_{ij} \neq 0 \mid y)$.


```{r}
run_simulations <- function(n, nsims, Theta, Sigma=solve(Theta), priorCoef, priorDiag, priorModel, scale=TRUE) {
  ans = vector("list", nsims)
  for (i in 1:nsims) {
    set.seed(i)  #for reproducibility

    ansi = vector("list", length(n))
    for (j in 1:length(n)) {
    
      # Simulate data
      samplesize <- n[j]
      y = rmvnorm(samplesize, sigma= Sigma)
      sy = apply(y, 2, 'sd')

      # Fit models   
      fit = modelSelectionGGM(y=y, priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel, center=scale, scale=scale, verbose=FALSE)

      # Extract BMA parameter estimates and posterior edge inclusion probabilities
      b= data.frame(coef(fit)) |>
        select(row, column, estimate, margpp)

      # Convert to original scale
      scaling = 1 / (sy[b$row] * sy[b$column])
      b = transform(b, estimate= estimate * scaling)

      # Add data-generating truth
      if (i==1) {
        truth = double(nrow(b))
        for (l in 1:nrow(b)) truth[l] = Theta[b$row[l], b$column[l]]
      }
      b$truth = truth
  
      # Record MSE, average posterior inclusion probability and average point estimate
      ansi[[j]] = group_by(b, truth) |>
                 summarize(mse = mean((truth - estimate)^2), pp = mean(margpp), mean_estimate = mean(estimate), count = n()) |>
                 cbind(simid = i, n = samplesize)

      cat('.')

    }
    
    ans[[i]] <- do.call(rbind, ansi)
  }
  return(ans)
}
```


# Banded diagonal

## Run simulations


```{r}
sims1 <- sims2 <- sims3 <- vector("list", length(kseq))
names(sims1) <- names(sims2) <- names(sims3) <- names(kseq)
diagvals <- vector("list", length(kseq))
for (i in 1:length(kseq)) {
  ksel <- kseq[i]  # number of variables
  cat("Simulations for k =", ksel)
  
  # Data-generating precision matrix
  Theta= diag(ksel)
  Theta[abs(col(Theta) - row(Theta)) == 1] <- 0.5
  Theta[abs(col(Theta) - row(Theta)) == 2] <- 0.25
  Sigma = solve(Theta)
  diagvals[[i]] <- unique(diag(Theta))

  # Set prior distribution under each setting
  pars_sel <- filter(priorpars, setting == "Setting 1", k == ksel)
  priorModel <- modelbinomprior(pars_sel$edge_prob)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims1[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel) 

  pars_sel <- filter(priorpars, setting == "Setting 2", k == ksel)
  priorModel <- modelbinomprior(pars_sel$edge_prob)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims2[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel) 

  pars_sel <- filter(priorpars, setting == "Setting 3", k == ksel)
  priorModel <- modelbinomprior(pars_sel$edge_prob)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims3[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel) 

  cat("\n")
}
diagvals <- unique(unlist(diagvals))
```


We summarize the results across all simulations. For each prior parameter setting and $k= \mbox{dim}(\Theta)$, and each value of the data-generating $\Theta_{ij}$, we report the mean squared estimation error in the Bayesian model averaging estimate $E(\Theta_{ij} \mid y)$, the average marginal posterior edge inclusion probability $P(\Theta_{ij} \neq 0 \mid y)$, and the average BMA estimate.

```{r}
summarize_simulations <- function(kseq, sims1, sims2, sims3) {
  
  sim_summary <- vector("list", length(kseq))
  names(sim_summary) <- names(kseq)

  for (i in 1:length(kseq)) {

    summary1 = do.call(rbind, sims1[[i]]) |>
      data.frame() |>
      group_by(n, truth) |>
      summarize(across(mse:count, mean), .groups="drop") |>
      transform(k = kseq[i], Setting = "Setting 1")

    summary2 = do.call(rbind, sims2[[i]]) |>
      data.frame() |>
      group_by(n, truth) |>
      summarize(across(mse:count, mean), .groups="drop") |>
      transform(k = kseq[i], Setting = "Setting 2")

    summary3 = do.call(rbind, sims3[[i]]) |>
      data.frame() |>
      group_by(n, truth) |>
      summarize(across(mse:count, mean), .groups="drop") |>
      transform(k = kseq[i], Setting = "Setting 3")
  
    sim_summary[[i]] <- rbind(summary1, summary2, summary3)

  }

  sim_summary <- do.call(rbind, sim_summary) |>
    transform(root_mse = round(sqrt(mse),3), pp = round(pp,3), mean_estimate = round(mean_estimate, 4)) |>
    select(-mse) |>
    select(Setting, k, root_mse, everything())

  rownames(sim_summary) <- NULL

  return(sim_summary)
}

```


```{r, warning=FALSE}
sim_summary <- summarize_simulations(kseq=kseq, sims1=sims1, sims2=sims2, sims3=sims3)
sim_summary
```


## Plots

We first display the root mean square error, and subsequently the average posterior inclusion probability.

```{r}
for (i in 1:length(kseq)) {
  
  myplot <- filter(sim_summary, !(truth %in% diagvals), k==kseq[i]) |>
  ggplot(aes(x=truth, y=root_mse, col=Setting, lty=Setting, pch=Setting)) +
    geom_line() +
    geom_point() +
    facet_wrap(~ n, labeller = labeller(n = function(x) paste0("n = ", x))) +
    theme_bw(base_size = 15) +
    theme(legend.position = c(0.1, 0.85), legend.title=element_blank()) +
    scale_color_grey() +
    xlab(expression("Data-generating " * Theta[ij])) +
    ylab("Root mean squared error")
  print(myplot)

  fname <- paste0("sim_priorsettings_banded_rootmse_nodiag_k", kseq[i], ".pdf")
  if (save2pdf) ggsave(fname, plot = myplot)
}
```

```{r}
for (i in 1:length(kseq)) {
  
  myplot <- filter(sim_summary, k==kseq[i]) |>
  ggplot(aes(x=truth, y=root_mse, col=Setting, lty=Setting, pch=Setting)) +
    geom_line() +
    geom_point() +
    facet_wrap(~ n, labeller = labeller(n = function(x) paste0("n = ", x))) +
    theme_bw(base_size = 15) +
    theme(legend.position = c(0.1, 0.85), legend.title=element_blank()) +
    scale_color_grey() +
    xlab(expression("Data-generating " * Theta[ij])) +
    ylab("Root mean squared error")
  print(myplot)

  fname <- paste0("sim_priorsettings_banded_rootmse_k", kseq[i], ".pdf")
  if (save2pdf) ggsave(fname, plot=myplot)
}
```

```{r}
for (i in 1:length(kseq)) {
  
  myplot <- filter(sim_summary, !(truth %in% diagvals), k==kseq[i]) |>
  ggplot(aes(x=truth, y=pp, col=Setting, lty=Setting, pch=Setting)) +
    geom_line() +
    geom_point() +
    facet_wrap(~ n, labeller = labeller(n = function(x) paste0("n = ", x))) +
    theme_bw(base_size = 15) +
    theme(legend.position = c(0.1, 0.85), legend.title=element_blank()) +
    scale_color_grey() +
    xlab(expression("Data-generating " * Theta[ij])) +
    ylab("Average posterior inclusion probability")
  print(myplot)

  fname <- paste0("sim_priorsettings_banded_pp_k", kseq[i], ".pdf")
  if (save2pdf) ggsave(fname, plot = myplot)
}
```


# Star graph

## Run simulations


```{r}
sims1 <- sims2 <- sims3 <- vector("list", length(kseq))
names(sims1) <- names(sims2) <- names(sims3) <- names(kseq)
diagvals <- vector("list", length(kseq))
for (i in 1:length(kseq)) {
  ksel <- kseq[i]  # number of variables
  cat("Simulations for k =", ksel)
  
  # Data-generating precision matrix
  Theta <- diag(ksel)
  Theta[1, 2:ksel] <- 0.5
  Theta[2:ksel, 1] <- 0.5
  diag(Theta) <- rowSums(abs(Theta)) + 0.01
  Sigma = solve(Theta)
  diagvals[[i]] <- unique(diag(Theta))

  # Set prior distribution under each setting
  pars_sel <- filter(priorpars, setting == "Setting 1", k == ksel)
  priorModel <- modelbinomprior(pars_sel$edge_prob)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims1[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel, scale=FALSE) 

  pars_sel <- filter(priorpars, setting == "Setting 2", k == ksel)
  priorModel <- modelbinomprior(pars_sel$edge_prob)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims2[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel, scale=FALSE) 

  pars_sel <- filter(priorpars, setting == "Setting 3", k == ksel)
  priorModel <- modelbinomprior(pars_sel$edge_prob)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims3[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel, scale=FALSE) 

  cat("\n")
}
diagvals <- unique(unlist(diagvals))
```


We summarize the results across all simulations. For each prior parameter setting and $k= \mbox{dim}(\Theta)$, and each value of the data-generating $\Theta_{ij}$, we report the mean squared estimation error in the Bayesian model averaging estimate $E(\Theta_{ij} \mid y)$, the average marginal posterior edge inclusion probability $P(\Theta_{ij} \neq 0 \mid y)$, and the average BMA estimate.

```{r, warning=FALSE}
sim_summary <- summarize_simulations(kseq=kseq, sims1=sims1, sims2=sims2, sims3=sims3)
sim_summary
```


## Plots

We first display the root mean square error, and subsequently the average posterior inclusion probability.

```{r}
for (i in 1:length(kseq)) {
  
  myplot <- filter(sim_summary, !(truth %in% diagvals), k==kseq[i]) |>
  ggplot(aes(x=truth, y=root_mse, col=Setting, lty=Setting, pch=Setting)) +
    geom_line() +
    geom_point() +
    facet_wrap(~ n, labeller = labeller(n = function(x) paste0("n = ", x))) +
    theme_bw(base_size = 15) +
    theme(legend.position = c(0.1, 0.85), legend.title=element_blank()) +
    scale_color_grey() +
    xlab(expression("Data-generating " * Theta[ij])) +
    ylab("Root mean squared error")
  print(myplot)

  fname <- paste0("sim_priorsettings_star_rootmse_nodiag_k", kseq[i], ".pdf")
  if (save2pdf) ggsave(fname, plot = myplot)
}
```

```{r}
for (i in 1:length(kseq)) {
  
  myplot <- filter(sim_summary, k==kseq[i]) |>
  ggplot(aes(x=truth, y=root_mse, col=Setting, lty=Setting, pch=Setting)) +
    geom_line() +
    geom_point() +
    facet_wrap(~ n, labeller = labeller(n = function(x) paste0("n = ", x))) +
    theme_bw(base_size = 15) +
    theme(legend.position = c(0.1, 0.85), legend.title=element_blank()) +
    scale_color_grey() +
    xlab(expression("Data-generating " * Theta[ij])) +
    ylab("Root mean squared error")
  print(myplot)

  fname <- paste0("sim_priorsettings_star_rootmse_k", kseq[i], ".pdf")
  if (save2pdf) ggsave(fname, plot = myplot)
}
```

```{r}
for (i in 1:length(kseq)) {
  
  myplot <- filter(sim_summary, !(truth %in% diagvals), k==kseq[i]) |>
  ggplot(aes(x=truth, y=pp, col=Setting, lty=Setting, pch=Setting)) +
    geom_line() +
    geom_point() +
    facet_wrap(~ n, labeller = labeller(n = function(x) paste0("n = ", x))) +
    theme_bw(base_size = 15) +
    theme(legend.position = c(0.1, 0.85), legend.title=element_blank()) +
    scale_color_grey() +
    xlab(expression("Data-generating " * Theta[ij])) +
    ylab("Average posterior inclusion probability")
  print(myplot)

  fname <- paste0("sim_priorsettings_star_pp_k", kseq[i], ".pdf")
  if (save2pdf) ggsave(fname, plot = myplot)
}
```



# Erdos-Renyi graph

## Run simulations


```{r}
sims1 <- sims2 <- sims3 <- vector("list", length(kseq))
names(sims1) <- names(sims2) <- names(sims3) <- names(kseq)
diagvals <- vector("list", length(kseq))
for (i in 1:length(kseq)) {
  ksel <- kseq[i]  # number of variables
  cat("Simulations for k =", ksel)
  
  # Data-generating precision matrix
  y <- matrix(nrow=0, ncol=ksel)
  sample <- modelSelectionGGM(y, priorCoef=normalidprior(1), priorModel=modelbinomprior(0.1), priorDiag=exponentialprior(1), niter=500, verbose=FALSE)
  Theta <- diag(ksel)
  thetavals <- sample$postSample[nrow(sample$postSample),]
  for (j in 1:ncol(sample$indexes)) {
    Theta[sample$indexes['row',j], sample$indexes['column',j]] <- thetavals[j]
    Theta[sample$indexes['column',j], sample$indexes['row',j]] <- thetavals[j]
  }
  Sigma <- solve(Theta)
  diagvals[[i]] <- unique(diag(Theta))

  # Set prior distribution under each setting
  pars_sel <- filter(priorpars, setting == "Setting 1", k == ksel)
  priorModel <- modelbinomprior(pars_sel$edge_prob)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims1[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel) 

  pars_sel <- filter(priorpars, setting == "Setting 2", k == ksel)
  priorModel <- modelbinomprior(pars_sel$edge_prob)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims2[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel) 

  pars_sel <- filter(priorpars, setting == "Setting 3", k == ksel)
  priorModel <- modelbinomprior(pars_sel$edge_prob)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims3[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel) 

  cat("\n")
}
diagvals <- unique(unlist(diagvals))
```


We summarize the results across all simulations. For each prior parameter setting and $k= \mbox{dim}(\Theta)$, and each value of the data-generating $\Theta_{ij}$, we report the mean squared estimation error in the Bayesian model averaging estimate $E(\Theta_{ij} \mid y)$, the average marginal posterior edge inclusion probability $P(\Theta_{ij} \neq 0 \mid y)$, and the average BMA estimate.

```{r, warning=FALSE}
sim_summary <- summarize_simulations(kseq=kseq, sims1=sims1, sims2=sims2, sims3=sims3) |>
  filter(!(truth %in% diagvals))

# To facilite plotting, group data-generating truth into bins
# (small bin around 0, to assess post prob to true zeroes)
breaks <- c(min(sim_summary$truth)-0.001,-2,-1,-0.01,0.01,1,2,max(sim_summary$truth)+0.001)
labels <- round((breaks[-1] + breaks[-length(breaks)])/2, 2)
sim_summary <- transform(sim_summary, truth = cut(truth, breaks=breaks, labels=labels)) |>
  group_by(Setting, k, n, truth) |>
  summarize(root_mse = mean(root_mse), pp = mean(pp), mean_estimate=mean(mean_estimate), .groups = "drop") |>
  transform(truth = as.numeric(as.character(truth)))
```


## Plots

We first display the root mean square error, and subsequently the average posterior inclusion probability.

```{r}
for (i in 1:length(kseq)) {
  
  myplot <- filter(sim_summary, !(truth %in% diagvals), k==kseq[i]) |>
  ggplot(aes(x=truth, y=root_mse, col=Setting, lty=Setting, pch=Setting)) +
    geom_line() +
    geom_point() +
    facet_wrap(~ n, labeller = labeller(n = function(x) paste0("n = ", x))) +
    theme_bw(base_size = 15) +
    theme(legend.position = c(0.1, 0.85), legend.title=element_blank()) +
    scale_color_grey() +
    xlab(expression("Data-generating " * Theta[ij])) +
    ylab("Root mean squared error")
  print(myplot)

  fname <- paste0("sim_priorsettings_erdosrenyi_rootmse_nodiag_k", kseq[i], ".pdf")
  if (save2pdf) ggsave(fname, plot = myplot)
}
```

```{r}
for (i in 1:length(kseq)) {
  
  myplot <- filter(sim_summary, k==kseq[i]) |>
  ggplot(aes(x=truth, y=root_mse, col=Setting, lty=Setting, pch=Setting)) +
    geom_line() +
    geom_point() +
    facet_wrap(~ n, labeller = labeller(n = function(x) paste0("n = ", x))) +
    theme_bw(base_size = 15) +
    theme(legend.position = c(0.1, 0.85), legend.title=element_blank()) +
    scale_color_grey() +
    xlab(expression("Data-generating " * Theta[ij])) +
    ylab("Root mean squared error")
  print(myplot)

  fname <- paste0("sim_priorsettings_erdosrenyi_rootmse_k", kseq[i], ".pdf")
  if (save2pdf) ggsave(fname, plot = myplot)
}
```

```{r}
for (i in 1:length(kseq)) {
  
  myplot <- filter(sim_summary, !(truth %in% diagvals), k==kseq[i]) |>
  ggplot(aes(x=truth, y=pp, col=Setting, lty=Setting, pch=Setting)) +
    geom_line() +
    geom_point() +
    facet_wrap(~ n, labeller = labeller(n = function(x) paste0("n = ", x))) +
    theme_bw(base_size = 15) +
    theme(legend.position = c(0.9, 0.15), legend.title=element_blank()) +
    scale_color_grey() +
    xlab(expression("Data-generating " * Theta[ij])) +
    ylab("Average posterior inclusion probability")
  print(myplot)

  fname <- paste0("sim_priorsettings_erdosrenyi_pp_k", kseq[i], ".pdf")
  if (save2pdf) ggsave(fname, plot = myplot)
}
```




# Block-diagonal graph

## Run simulations


```{r}
sims1 <- sims2 <- sims3 <- vector("list", length(kseq))
names(sims1) <- names(sims2) <- names(sims3) <- names(kseq)
diagvals <- vector("list", length(kseq))
for (i in 1:length(kseq)) {
  ksel <- kseq[i]  # number of variables
  cat("Simulations for k =", ksel)
  
  # Data-generating precision matrix
  block_size <- 5
  num_blocks <- ksel / block_size
  B <- matrix(0.5, nrow = block_size, ncol = block_size)
  diag(B) <- 1
  Theta <- kronecker(diag(num_blocks), B)
  Sigma <- solve(Theta)
  diagvals[[i]] <- unique(diag(Theta))

  # Set prior distribution under each setting
  pars_sel <- filter(priorpars, setting == "Setting 1", k == ksel)
  priorModel <- modelbinomprior(pars_sel$edge_prob)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims1[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel) 

  pars_sel <- filter(priorpars, setting == "Setting 2", k == ksel)
  priorModel <- modelbinomprior(pars_sel$edge_prob)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims2[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel) 

  pars_sel <- filter(priorpars, setting == "Setting 3", k == ksel)
  priorModel <- modelbinomprior(pars_sel$edge_prob)
  priorDiag <- exponentialprior(lambda = pars_sel$lambda)
  priorCoef <- normalidprior(pars_sel$var.slab)
  sims3[[i]] <- run_simulations(n=nseq, nsims=nsims, Theta=Theta, Sigma=solve(Theta), priorCoef=priorCoef, priorDiag=priorDiag, priorModel=priorModel) 

  cat("\n")
}
diagvals <- unique(unlist(diagvals))
```


We summarize the results across all simulations. For each prior parameter setting and $k= \mbox{dim}(\Theta)$, and each value of the data-generating $\Theta_{ij}$, we report the mean squared estimation error in the Bayesian model averaging estimate $E(\Theta_{ij} \mid y)$, the average marginal posterior edge inclusion probability $P(\Theta_{ij} \neq 0 \mid y)$, and the average BMA estimate.

```{r, warning=FALSE}
sim_summary <- summarize_simulations(kseq=kseq, sims1=sims1, sims2=sims2, sims3=sims3)
sim_summary
```


## Plots

We first display the root mean square error, and subsequently the average posterior inclusion probability.

```{r}
for (i in 1:length(kseq)) {
  
  myplot <- filter(sim_summary, !(truth %in% diagvals), k==kseq[i]) |>
  ggplot(aes(x=truth, y=root_mse, col=Setting, lty=Setting, pch=Setting)) +
    geom_line() +
    geom_point() +
    facet_wrap(~ n, labeller = labeller(n = function(x) paste0("n = ", x))) +
    theme_bw(base_size = 15) +
    theme(legend.position = c(0.1, 0.85), legend.title=element_blank()) +
    scale_color_grey() +
    xlab(expression("Data-generating " * Theta[ij])) +
    ylab("Root mean squared error")
  print(myplot)

  fname <- paste0("sim_priorsettings_block_rootmse_nodiag_k", kseq[i], ".pdf")
  if (save2pdf) ggsave(fname, plot = myplot)
}
```

```{r}
for (i in 1:length(kseq)) {
  
  myplot <- filter(sim_summary, k==kseq[i]) |>
  ggplot(aes(x=truth, y=root_mse, col=Setting, lty=Setting, pch=Setting)) +
    geom_line() +
    geom_point() +
    facet_wrap(~ n, labeller = labeller(n = function(x) paste0("n = ", x))) +
    theme_bw(base_size = 15) +
    theme(legend.position = c(0.1, 0.85), legend.title=element_blank()) +
    scale_color_grey() +
    xlab(expression("Data-generating " * Theta[ij])) +
    ylab("Root mean squared error")
  print(myplot)

  fname <- paste0("sim_priorsettings_block_rootmse_k", kseq[i], ".pdf")
  if (save2pdf) ggsave(fname, plot = myplot)
}
```

```{r}
for (i in 1:length(kseq)) {
  
  myplot <- filter(sim_summary, !(truth %in% diagvals), k==kseq[i]) |>
  ggplot(aes(x=truth, y=pp, col=Setting, lty=Setting, pch=Setting)) +
    geom_line() +
    geom_point() +
    facet_wrap(~ n, labeller = labeller(n = function(x) paste0("n = ", x))) +
    theme_bw(base_size = 15) +
    theme(legend.position = c(0.1, 0.85), legend.title=element_blank()) +
    scale_color_grey() +
    xlab(expression("Data-generating " * Theta[ij])) +
    ylab("Average posterior inclusion probability")
  print(myplot)

  fname <- paste0("sim_priorsettings_block_pp_k", kseq[i], ".pdf")
  if (save2pdf) ggsave(fname, plot = myplot)
}
```

