---
title: "Effect of +def truncation on separable prior"
output: 
  html_document:
    toc: true
    theme: united
---

```{r, warning=FALSE, message=FALSE}
library(modelSelection)
library(tidyverse)
library(lamW)      # Lambert's W_q function
set.seed(20260606) # for reproducibility
save2pdf <- FALSE  # set to true to write plots to pdf
```

# Set simulation parameters

We consider two problem dimensions $k \in \{5, 50\}$ and four prior edge inclusion probabilities $\eta \in \{ 0.05, 0.1, 0.2, 0.4 \}$.
Non-zero off-diagonal entries are simulated $\Theta_{ij} \mid Z_{ij}=1 \sim N(0, \sigma^2)$ where $\sigma^2=1$, and $Z_{ij}= \mbox{I}(\Theta_{ij} \neq 0)$ is an edge inclusion indicator for $i > j$.
Diagonal entries are generated $\Theta_{ii} \sim \mbox{Exp}(\lambda / 2)$, where $\lambda / 2$ is the rate parameter and hence $2/\lambda$ the mean. As a first possible choice we consider $\lambda = 1$.

```{r}
kseq <- c(5, 50)
edge_prob_seq <- c(0.05, 0.1, 0.2, 0.4)
lambda1 <- 1
```

As a second possible choice for $\lambda$, we set $\lambda$ such that $P(\Theta_{ii} \succ 0 \mid Z) \geq 1 - \alpha$, for every edge configuration $Z$, taking $\alpha=0.1$. Using the formulae given in the main paper, this is given by taking
$$
\frac{2}{\lambda} > \frac{2 \sigma p \sqrt{\bar d}}{\alpha} \sqrt{- W_{-1}\left( \frac{-\alpha^2}{32 e^2 p^2} \right)},
$$
where $W_{-1}$ is the branch of Lambert's $W_q$ function associated to $q=-1$, and $\bar d$ is an upper-bound on the maximum edge degree. Below we show results for $\bar d =1$, which results in a somewhat liberal upper bound, but virtually identical results are obtained by setting $\bar d = p$. Note that this rule gives a value of $\lambda$ that depends on $p$, here for simplicity we take the $\lambda$ value associated to the larger $k=100$, since this also suffices for any $p \leq 100$ (and the results remain virtually identical when using a different $\lambda$ for $k=5$).

```{r}
sd.slab <- 1
alpha <- 0.1
p <- max(kseq)
x <- - alpha^2 /(32 * exp(2) * p^2)
l <- lambertWm1(x) # q = -1 branch of Lambert's W_q function
mu <- (2 * sd.slab * p / alpha) * sqrt(-l)  # liberal upper bound
lambda2 <- 2/mu
#mu <- (2 * sd.slab * p^(3/2) / alpha) * sqrt(-l) # conservative upper-bound (same results)
lambda2
```


# Simulate from prior

```{r}
fit1 <- fit2 <- lapply(1:length(kseq), function(z) vector("list", length(edge_prob_seq)))
priorCoef <- normalidprior(sd.slab^2)
for (i in 1:length(kseq)) {
  p <- kseq[i]
  y <- matrix(nrow=0, ncol=p)
  for (j in 1:length(edge_prob_seq)) {
    priorModel <- modelbinomprior(edge_prob_seq[j])

    fit1[[i]][[j]] <- modelSelectionGGM(y, niter=5000, burnin=0, priorCoef=priorCoef, priorModel=priorModel, 
                                        priorDiag= exponentialprior(lambda1), verbose=FALSE)
    fit2[[i]][[j]] <- modelSelectionGGM(y, niter=5000, burnin=0, priorCoef=priorCoef, priorModel=priorModel, 
                                        priorDiag= exponentialprior(lambda2), verbose=FALSE)
    cat('.')
  }
  cat('\n')
}
```

Assess MCMC convergence. A burnin of 1,000 is enough

```{r}
par(mar= c(5,5,0.1,0.1), mfrow=c(2,1))
for (i in 1:length(kseq)) {
  for (j in 1:length(edge_prob_seq)) {
    plot(fit1[[i]][[j]]) 
    text(5000, 0.95, paste('k=', kseq[i], sep=''), pos=2)
    text(5000, 0.80, paste('edge_prob=', edge_prob_seq[j], sep=''), pos=2)
    text(5000, 0.65, paste('lambda=', lambda1, sep=''), pos=2)
    plot(fit2[[i]][[j]])
    text(5000, 0.95, paste('k=', kseq[i], sep=''), pos=2)
    text(5000, 0.80, paste('edge_prob=', edge_prob_seq[j], sep=''), pos=2)
    text(5000, 0.65, paste('lambda=', lambda2, sep=''), pos=2)
  }
}
```

# Format simulation output

Store off-diagonal samples in a vector, discarding the burnin.

```{r}
burnin <- 1000

empty_named_list <- function(z, element_names) {
  ans <- vector("list", length(edge_prob_seq))
  names(ans) <- element_names
  return(ans)
}

element_names <- paste("edge_prob = ", edge_prob_seq, sep='')
diagsamples1 <- lapply(1:length(kseq), empty_named_list, element_names=element_names)
names(diagsamples1) <- paste("k = ", kseq, sep='')
diagsamples2 <- offdiagsamples1 <- offdiagsamples2 <- total_edges1 <- total_edges2 <- diagsamples1


for (i in 1:length(kseq)) {
  for (j in 1:length(edge_prob_seq)) {
    diagentry <- fit1[[i]][[j]]$indexes[1,] == fit1[[i]][[j]]$indexes[2,]
    diagsamples1[[i]][[j]] <- as.vector(fit1[[i]][[j]]$postSample[-1:-burnin,diagentry])
    diagsamples2[[i]][[j]] <- as.vector(fit2[[i]][[j]]$postSample[-1:-burnin,diagentry])
    offdiagsamples1[[i]][[j]] <- as.vector(fit1[[i]][[j]]$postSample[-1:-burnin,!diagentry])
    offdiagsamples2[[i]][[j]] <- as.vector(fit2[[i]][[j]]$postSample[-1:-burnin,!diagentry])
    total_edges1[[i]][[j]] <- rowSums(as.matrix(fit1[[i]][[j]]$postSample[-1:-burnin,!diagentry] != 0))
    total_edges2[[i]][[j]] <- rowSums(as.matrix(fit2[[i]][[j]]$postSample[-1:-burnin,!diagentry] != 0))
  }
}
```

# Extract results

For each $p$ and edge probability, we calculate the prior edge probability after truncation, standard deviation of non-zero off-diagonal entries, and mean diagonal entry.

```{r}
pars <- expand.grid(k=factor(kseq), edge_prob=edge_prob_seq)

edge_prob_trunc <- sd_offdiag <- mean_diag <- matrix(NA, nrow=nrow(pars), ncol=2)
colnames(edge_prob_trunc) <- paste("lambda=",round(c(lambda1,lambda2),5))
sd_nonzero <- function(x) sd(x[x != 0])
for (i in 1:nrow(pars)) {
  k <- pars$k[i]
  edge_prob <- pars$edge_prob[i]
  
  sel1 <- which(names(diagsamples1) == paste('k = ', k, sep=''))
  sel2 <- which(names(diagsamples1[[1]]) == paste('edge_prob = ', edge_prob, sep=''))
  
  edge_prob_trunc[i,1] <- mean(offdiagsamples1[[sel1]][[sel2]] != 0) 
  edge_prob_trunc[i,2] <- mean(offdiagsamples2[[sel1]][[sel2]] != 0) 
  sd_offdiag[i, 1] <- sd_nonzero(offdiagsamples1[[sel1]][[sel2]])
  sd_offdiag[i, 2] <- sd_nonzero(offdiagsamples2[[sel1]][[sel2]])
  mean_diag[i, 1] <- mean(diagsamples1[[sel1]][[sel2]])
  mean_diag[i, 2] <- mean(diagsamples2[[sel1]][[sel2]])
}
```

```{r}
edge_prob_trunc <- cbind(pars, edge_prob_trunc) |>
  pivot_longer(cols=3:4, names_to="lambda") |>
  mutate(value = round(value, 3)) |>
  rename(edge_prob_trunc = value) |>
  mutate(lambda = factor(lambda)) |>
  mutate(lambda = factor(lambda, levels=rev(levels(lambda))))

sd_offdiag <- cbind(pars, sd_offdiag) |>
  pivot_longer(cols=3:4, names_to="lambda") |>
  mutate(value = round(value, 3)) |>
  rename(sd_offdiag = value)

mean_diag <- cbind(pars, mean_diag) |>
  pivot_longer(cols=3:4, names_to="lambda") |>
  mutate(value = round(value, 3)) |>
  rename(mean_diag = value)

```

We also obtain the prior distribution of the number of edges, before and after truncation (the former is given by a Binomial distribution, the latter is estimated from the MCMC output).

```{r}
nedges1 <- lapply(total_edges1, function(z) data.frame(do.call(cbind, z))) |>
  lapply(pivot_longer, cols=1:length(edge_prob_seq), names_to="edge_prob")
nedges1 <- do.call(cbind, nedges1) |>
  select(-3)
names(nedges1) <- sub(".value","",names(nedges1))
names(nedges1)[1] <- "edge_prob"
nedges1 <- mutate(nedges1, edge_prob = sub("edge_prob...","",edge_prob)) |>
  pivot_longer(cols=2:3, names_to= "k") |>
  mutate(k = sub("k = ","", k)) |>
  rename(number_edges = value)

nedges2 <- lapply(total_edges2, function(z) data.frame(do.call(cbind, z))) |>
  lapply(pivot_longer, cols=1:length(edge_prob_seq), names_to="edge_prob")
nedges2 <- do.call(cbind, nedges2) |>
  select(-3)
names(nedges2) <- sub(".value","",names(nedges2))
names(nedges2)[1] <- "edge_prob"
nedges2 <- mutate(nedges2, edge_prob = sub("edge_prob...","",edge_prob)) |>
  pivot_longer(cols=2:3, names_to= "k") |>
  mutate(k = sub("k = ","", k)) |>
  rename(number_edges = value)
```

We only evaluate the Binomial prior for $k=100$, since we will later restrict the plot to this case.

```{r}
totaledges <- max(kseq) * (max(kseq)-1) / 2
nedges_binom <- vector("list", length(edge_prob_seq))
for (j in 1:length(edge_prob_seq)) {
  
  pmf_edges <- dbinom(x=0:totaledges, size=totaledges, prob=edge_prob_seq[j])
  number_edges <- rep(0:totaledges, round(10000 * pmf_edges, 4))
  
  nedges_binom[[j]] <- data.frame(edge_prob=edge_prob_seq[j],
                                  number_edges = number_edges) |>
                       mutate(trunc = "Before truncation")
  
}
nedges_binom <- do.call(rbind, nedges_binom)
```


# Visualize results

First we display the computed quantities, and subsequently we plot them.

```{r}
edge_prob_trunc
```

```{r}
sd_offdiag
```

```{r}
mean_diag
```

## Edge probabilities

We plot the edge inclusion probabilities before and after truncation.

```{r}
ggplot(edge_prob_trunc, aes(x=edge_prob, y=edge_prob_trunc, col=k)) +
  geom_line() +
  geom_abline(lty=2) +
  facet_wrap(~ lambda) +
  theme_bw(base_size = 15) +
  theme(legend.position = c(0.9, 0.2)) +
  scale_color_grey() +
  xlab("Edge probability (before truncation)") +
  ylab("Edge probability (after truncation)")

if (save2pdf) ggsave("prior_edge_prob.pdf")
```

We also plot the prior distribution on the number of edges.

```{r}
data2plot <- filter(nedges1, k=="50") |>
  select(-k) |>
  mutate(edge_prob = as.numeric(edge_prob), trunc = "After truncation") |>
  rbind(nedges_binom) |>
  mutate(edge_prob = paste("Edge probability = ", edge_prob, sep=''))
```


```{r, warning=FALSE, message=FALSE}
ggplot(data2plot, aes(x=number_edges, col=trunc)) +
  geom_histogram(aes(y = after_stat(density)), fill=NA, position="identity") +
  facet_wrap(~ edge_prob, scales="free_x") +
  theme_bw(base_size = 15) +
  theme(legend.position = c(0.9, 0.3), legend.title = element_blank()) +
  scale_color_grey() +
  xlab("Number of edges") +
  ylab("Prior probability")
```


We zoom into the plots for edge probabilities 0.05 and 0.2.

```{r}
filter(data2plot, edge_prob == "Edge probability = 0.05") |>
ggplot(aes(x=number_edges, col=trunc)) +
  geom_histogram(aes(y = after_stat(density)), fill=NA, position="identity") +
  theme_bw(base_size = 15) +
  theme(legend.position = c(0.85, 0.85), legend.title = element_blank()) +
  scale_color_grey() +
  xlab("Number of edges") +
  ylab("Prior probability")

if (save2pdf) ggsave("prior_number_edges_edgeprob05.pdf")

filter(data2plot, edge_prob == "Edge probability = 0.2") |>
ggplot(aes(x=number_edges, col=trunc)) +
  geom_histogram(aes(y = after_stat(density)), fill=NA, position="identity") +
  theme_bw(base_size = 15) +
  theme(legend.position = c(0.85, 0.85), legend.title = element_blank()) +
  scale_color_grey() +
  xlab("Number of edges") +
  ylab("Prior probability")

if (save2pdf) ggsave("prior_number_edges_edgeprob20.pdf")

```


## Diagonal entries

Plot density before and after positive-definiteness truncation. First, for diagonal entries. We create a function that obtains a kernel density estimate, via taking a log transformation of the input and then performing the associated change of variables.

```{r}
log_density <- function(x, threshold=0.1, ...) {
  if (any(x <= 0, na.rm = TRUE)) stop("All values of x must be > 0")

  y <- log(x)
  dens_y <- density(y, ...)
  x_grid <- exp(dens_y$x)
  
  sel <- (x_grid > threshold)
  x_grid <- x_grid[sel] # avoid numerical instability close to 0
  f_x <- dens_y$y[sel] / x_grid   # apply Jacobian term (divide by x)

  ans <- list(x = x_grid, y = f_x, bw = dens_y$bw, n = dens_y$n, call = match.call(),
      data.name = deparse(substitute(x)), has.na = dens_y$has.na)
  structure(ans, class = "density") # return a 'density' object
}
```

```{r, fig.show = "hold", out.width = "50%"}
col <- c('gray', 'black', 'black')
lwd <- c(2,rep(1,length(kseq)))
lty <- c(1, 1:length(kseq))
for (j in 1:length(edge_prob_seq)) {

  edge_prob <- edge_prob_seq[j]
  ylab <- paste('Prior density (edge prob=', edge_prob,')',sep='')
  lab <- c('Before truncation', paste('After truncation (k=', kseq, ')',sep=''))

  # Plot for lambda = 1
  dseq <- seq(0.01, max(diagsamples1[[1]][[j]]), length=200)
  fname <- paste("priordiag_edgeprob_", edge_prob_seq[j], "_lambda1.pdf", sep='')
  if (save2pdf) pdf(fname)
  par(mar= c(5,5,0.1,0.1), cex.lab=1.5, cex.axis=1.5, mfrow=c(1,1))

  plot(dseq, dexp(dseq, rate=lambda1/2), xlab="Diagonal entry", ylab=ylab, 
       type='l', col=col[1], lwd=lwd[1], lty=lty[1])
  for (i in 1:length(kseq)) {
    lines(log_density(diagsamples1[[i]][[j]]), col=col[i+1], lty=lty[i+1], lwd=lwd[i+1])
  }
  
  legend('topright', lab, col= col, lwd=lwd, lty=lty, cex=1.3)

  if (save2pdf) dev.off()

  # Plot for smaller lambda
  dseq <- seq(0.01, qexp(0.99, rate=lambda2/2), length=200)
  fname <- paste("priordiag_edgeprob_", edge_prob_seq[j], "_lambda2.pdf", sep='')
  if (save2pdf) pdf(fname)
  par(mar= c(5,5,0.1,0.1), cex.lab=1.5, cex.axis=1.5, mfrow=c(1,1))

  plot(dseq, dexp(dseq, rate=lambda2/2), xlab="Diagonal entry", ylab=ylab, 
       type='l', col=col[1], lwd=lwd[1], lty=lty[1])
  for (i in 1:length(kseq)) {
    lines(log_density(diagsamples2[[i]][[j]]), col=col[i+1], lty=lty[i+1], lwd=lwd[i+1])
  }
  
  legend('topright', lab, col= col, lwd=lwd, lty=lty, cex=1.3)

  if (save2pdf) dev.off()

}

```


## Off-diagonal entries

Finally, we plot the marginal densities of the off-diagonal entries.

```{r, fig.show = "hold", out.width = "50%"}
col <- c('gray', 'black', 'black')
lwd <- c(2,rep(1,length(kseq)))
lty <- c(3, 1:length(kseq))
for (j in 1:length(edge_prob_seq)) {

  edge_prob <- edge_prob_seq[j]
  ylab <- paste('Prior density (edge prob=', edge_prob,')',sep='')
  lab <- c('Before truncation', paste('After truncation (k=', kseq, ')',sep=''))
  oseq <- seq(-3.5, 3.5, length=200)

  # Plot for lambda = 1
  fname <- paste("prioroffdiag_edgeprob_", edge_prob_seq[j], "_lambda1.pdf", sep='')
  if (save2pdf) pdf(fname)
  par(mar= c(5,5,0.1,0.1), cex.lab=1.5, cex.axis=1.5, mfrow=c(1,1))

  dnorm_oseq <- dnorm(oseq, sd=1)
  ylim <- c(0, 1.5*max(dnorm_oseq))
  plot(oseq, dnorm_oseq, xlab="Off-diagonal entry", ylab=ylab, 
       type='l', col=col[1], lwd=lwd[1], lty=lty[1], ylim=ylim)
  for (i in 1:length(kseq)) {
    nonzeroes <- offdiagsamples1[[i]][[j]]
    nonzeroes <- nonzeroes[nonzeroes != 0]
    lines(density(nonzeroes), col=col[i+1], lty=lty[i+1], lwd=lwd[i+1])
  }
  
  legend('topright', lab, col= col, lwd=lwd, lty=lty, cex=1.2)

  if (save2pdf) dev.off()

  # Plot for smaller lambda
  fname <- paste("prioroffdiag_edgeprob_", edge_prob_seq[j], "_lambda2.pdf", sep='')
  if (save2pdf) pdf(fname)
  par(mar= c(5,5,0.1,0.1), cex.lab=1.5, cex.axis=1.5, mfrow=c(1,1))

  plot(oseq, dnorm(oseq, sd=1), xlab="Off-diagonal entry", ylab=ylab, 
       type='l', col=col[1], lwd=lwd[1], lty=lty[1], ylim=ylim)
  for (i in 1:length(kseq)) {
    nonzeroes <- offdiagsamples2[[i]][[j]]
    nonzeroes <- nonzeroes[nonzeroes != 0]
    lines(density(nonzeroes), col=col[i+1], lty=lty[i+1], lwd=lwd[i+1])
  }
  
  legend('topright', lab, col= col, lwd=lwd, lty=lty, cex=1.2)

  if (save2pdf) dev.off()

}
```




